### 2017-05-10 conclusion.tex
Add files via upload

\chapter{Concluding statements}

This study presents the first long-term, time-continuous analysis of urban T\textsubscript{surf} to contribute to a more complete understanding of the temporal and geometric nature of the urban effect on land T\textsubscript{surf} and sUHI.

Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive atmospherically correct hemispherical radiometric surface temperatures from upwelling longwave radiation measured from a downward facing pyrgeometer. The correction method was developed to be readily adaptable to other sensor types and study sites, these measurements are commonly made as a part of urban energy balance assessment. By accounting for surface geometry and spectrally non-uniform sensor response, the correction method is readily adaptable to other sensor types and study sites. Atmospheric effects on inferred T\textsubscript{hem} are significant relative to typical sUHI magnitudes and show significant diurnal and seasonal variation. Correction magnitudes are particularly sensitive to $\Delta$T\textsubscript{surf-air}. Although water vapor is the main absorber of broadband TIR radiation, it does not exert significant influence on correction magnitudes, as atmospheric transmittance shows little variation at typical humidities.

In addition, Chapter \ref{paper1} explores the effect of surface geometry on remote sensed urban T\textsubscript{surf} for three representations of the urban surface (complete, plan, and hemispherical). Analysis shows significant overestimation (underestimation) of T\textsubscript{comp} by T\textsubscript{plan} by day (night), particularly under clear sky conditions. By day, T\textsubscript{hem} overestimates T\textsubscript{comp}, but to a lesser degree than T\textsubscript{plan}. Daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem} is more consistent under a wide variety of synoptic conditions, whereas overestimation by T\textsubscript{plan} fluctuates significantly, resulting in poor performance under clear-sky "satellite friendly" conditions. As make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warming bias. These biases are reflected in sUHI analysis in \ref{paper2}.

Analysis of an eight month, time continuous climatology of hemispherical sUHI is included in Chapter \ref{paper2}.

>>>>

\chapter{Conclusion}

This study presents the first long-term, time-continuous analysis of urban T\textsubscript{surf} and sUHI that incorporates representative sampling of the three-dimensional urban surface. It is based on hemispheric upwelling longwave radiation measurements that are commonly made as a part of urban energy balance assessment and constitute a hitherto untapped resource for urban T\textsubscript{surf} and sUHI analysis. The analysis contributes to a more complete understanding of the temporal and geometric nature of the urban effect on land T\textsubscript{surf} and sUHI.

\section{Summary of results}

Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive hemispherical radiometric surface temperatures from longwave radiation upwelling from the urban surface measured from a downward facing pyrgeometer. The correction method is necessary for two reasons:

\begin{enumerate}
	\item To retrieve T\textsubscript{surf} that is representative of the temperature at which the surface is radiating (known as radiometric T\textsubscript{surf}) by removing effects from absorption and emission by the atmosphere in between the surface and the sensor.
	\item To facilitate intersite analysis of T\textsubscript{surf}, where different ambient conditions, surface-sensor-sun geometries, and instrument types can yield differential atmospheric effects on the remote sensed TIR signal.
\end{enumerate}

 The correction method was developed to be readily adaptable to other sensor types and study sites by accounting for complex three-dimensional surface geometry and non uniform spectral sensor response.

Analysis of an eight months of corrected T\textsubscript{hem, r} shows that atmospheric effects on inferred T\textsubscript{hem} are large relative to typical sUHI magnitudes and show significant diurnal and seasonal variation. Relationships between correction magnitude and meteorological variables are complex. Correction magnitudes show strong positive relationships with $\Delta$T\textsubscript{surf-air} and incoming solar radiation. Conversely, although water vapor is the main absorber of broadband TIR radiation, it does not exert significant influence on correction magnitudes, as atmospheric transmittance shows little variation at typical humidities. To facilitate application of the method, a parameterization was developed incorporating a small number of generalizable radiative transfer simulations (included in Appendix \ref{appendixa}) and information about site geometry to retrieve radiometric hemispherical T\textsubscript{surf} en masse using measurements of upwelling longwave, humidity, and T\textsubscript{air}.

In addition, Chapter \ref{paper1} explores the effect of surface geometry on remote sensed urban T\textsubscript{surf} for three representations of the urban surface (complete, plan, and hemispherical). Analysis shows significant overestimation (underestimation) of T\textsubscript{comp} by T\textsubscript{plan} by day (night), particularly under clear sky conditions. By day, T\textsubscript{hem} overestimates T\textsubscript{comp}, but to a lesser degree than T\textsubscript{plan}. Daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem} is more consistent under a wide variety of synoptic conditions, whereas overestimation by T\textsubscript{plan} fluctuates significantly, resulting in poor performance under clear-sky "satellite friendly" conditions. As these measures make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warm bias. These biases are reflected in sUHI analysis in Chapter \ref{paper2}.

Chapter \ref{paper2} includes an analysis of two long term, time continuous climatologies of hemispherical sUHI, one in Basel, Switzerland and another in Vancouver, Canada. sUHI for both study sites shows significant diurnal, day-to-day, and seasonal variability. sUHI is highly dependent on meteorological conditions, particularly accumulated solar radiation over a day and wind velocity. However, as sUHI is the result of differences in heating and cooling rates between the city and its non urban surroundings, its relationship to meteorological controls is complex. Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of the first observationally supported generalizable model for summertime sUHI development presented in section \ref{gen}.

Analysis of the directional dependence of sUHI is also included in Chapter \ref{paper2}. Results are similar to those found in Chapter \ref{paper1}, with significant daytime overestimation of sUHI\textsubscript{comp} by sUHI{plan}, particularly during hot clear sky conditions. Not only did directional biases in remote sensed sUHI change observed magnitudes, these biases also changed the diurnal character of sUHI, modifying patterns of sUHI in the hours during near sunset. Thus, satellite based assessments of the diurnal pattern of sUHI may lead to misrepresentations of its 'true' diurnal character.

\section{Limitations and future work}

%The convoluted and complex nature of the urban surface makes remote sensing urban T\textsubscript{surf} difficult. Although the method detailed in this study provides T\textsubscript{surf} that is time continuous and more representative of the complete urban T\textsubscript{surf} than conventional remote sensing methods - particularly under clear sky conditions - it is not without some limitations.

The complex nature of the urban surface introduces myriad shortcomings and biases into conventional remote sensing of urban thermal climates - this method is no exception. Sensor placement sensitivity tests in Section \ref{placement} corroborate findings in \citet{Roberts2010,Adderley2015} and indicate that longwave radiation upwelling from the urban surface as measured from a downward facing radiometer is spatially variant when measured from heights below approximately three to five times mean building height. Most pyrgeometers (including those used in this study) are mounted below this height and are subject to a bias towards facets that make up a disproportionately large fraction of the sensor FOV. For sensors mounted at less than optimal heights, the effect of this bias on derived T\textsubscript{hem, r} is largely dependent on sensor placement relative to rooftop facets, with better performance further from positions over rooftops. Pyrgeometers at the Sperrstrasse and Sunset tower sites are placed far enough from rooftops to render this bias relatively small. However, future studies should strive for pyrgeometer heights that are more optimal not only to yield more geometrically representative T\textsubscript{hem, r}, but also to ensure irradiances are representative of the complete urban longwave emission in energy balance assessments.

As the method is fixed and ground based, it is not practical for large scale spatially extensive analysis of sUHI. This limits the method to mid- and low-rise neighborhoods. However, the concept can be adapted to retrieve spatiotemporally continuous urban T\textsubscript{hem, r} from mobile measurements of upwelling longwave irradiances. \citet{Sugawara2006} used near-ground, helicopter-mounted, pyrgeometers and narrow to quantify effects from urban effective anisotropy on upwelling longwave radiation, but did not incorporate a robust atmospheric correction routine to retrieve T\textsubscript{hem, r} and remove spatiotemporally variant atmospheric effects. Inherent complexity in determining the distance to the surface for a moving sensor "seeing" a complex, three-dimensional surface make sensor view modeling difficult. Ray-tracing methods similar to those used in \citet{Ceamanos2017} could provide a solution, but must be modified to return path lengths in addition to seen and unseen points. In addition, sensitivity to ambient conditions (particularly the T\textsubscript{air} profile) will make accurate atmospheric correction computationally expensive in mobile applications of the method. Notwithstanding evaluations of method performance in such applications. If these problems are solved, mobile application of the method may be useful for determine as hemispherical instrumentation can be run simultaneously with near-ground and satellite based narrow-FOV to further understanding of effective anisotropy on TIR remote sensing of urban areas. Such analysis could be used to develop parameterizations to modify satellite retrieved urban T\textsubscript{surf} to better represent the complete urban T\textsubscript{surf}.

Currently, little is known about. The simple fact that issues raised in have yet to be addressed should prompt diverse assessment of urban T\textsubscript{surf} via a range of sensors operating at wide range of spatiotemporal scales is necessary to foster better understand the true geometric, temporal, and spatial effects of urban areas on land T\textsubscript{surf}. While satellite assessment of sUHI and urban T\textsubscript{surf} has been instrumental in elucidating the spatial character of urban effects on T\textsubscript{surf}, several of the questions posed nearly 30 years ago in \citet{Roth1989} remain unanswered. This should prompt not only critical assessment of the urban T\textsubscript{surf} record, but also the development and use of novel methods to fill knowledge gaps, quantify biases, and uncover methodological shortcomings. This method, and in particular its parameterization, is intended to aid in such an endeavor. According to the Urban Flux Network (\href{http://fluxnet.fluxdata.org/}{Urbanflux}) there are 26 active urban flux towers, with many other existing inactive towers, in addition to datasets not indexed under the Urbanflux umbrella for which this method can be used for urban T\textsubscript{surf} and sUHI analysis. These span several continents, a range of climate zones and socioeconomic strata, with many operational over very long (5+ years) time scales. The application of this method over these datasets, and as a component of future energy balance assessments, could provide a wealth of potentially important analysis within and outside the purview of urban climatology.

\section{Final remarks}

%Broadly, the method, its parameterization, and the analysis presented in this study not only provides for novel analysis of the urban effect on T\textsubscript{surf}, but also aims to prompt further critical assessment of the urban T\textsubscript{surf} and the sUHI record.

Broadly, the method, its parameterization, and the analysis presented in this study was conducted to provide for novel analyses of the urban effect on T\textsubscript{surf} and to prompt further critical assessment of the urban T\textsubscript{surf} and sUHI record.  work towards addressing methodological and conceptual questions outlined in \citet{Roth1989}. Indeed, if similar conceptual flaws and shortcomings to those found for the aUHI in \citet{Stewart2011} are likely in sUHI analysis, quantification of instrument and methodological biases and shortcomings in remote sensing of urban T\textsubscript{surf} and the introduction of new methods to overcome these shortcomings may bring us closer to uncovering the true geometric, temporal, and spatial effects of cities on land surface temperature and on the climate at large.

### 2017-05-10 introduction.tex
Add files via upload

\chapter{Introduction}
Urban development drastically changes the character of the surface. Replacement of natural terrain with roads, buildings, parks, and other urban features modifies the geometry of the surface as well as its thermal, radiative, moisture, and aerodynamic properties. These modifications, combined with anthropogenic emissions of heat, result in a distinct urban climate; one which is often warmer, a phenomenon termed the urban heat island effect (UHI). As both the Earth's population and the urbanized fraction of that population increase \citep{Nations2014}, cities grow and more people are exposed to urban-modified atmospheres. Thus, understanding how cities interact with the climate across spatiotemporal scales has important implications for human health, energy efficiency, and for informing future urban development. This, in part, has prompted significant expansion in study of the urban effect on climates, with particular focus on characterization of the urban effect on surface (T\textsubscript{surf}) and air (T\textsubscript{air}) temperatures.

@@

Study of the UHI with respect to T\textsubscript{air} (aUHI) - typically referring to T\textsubscript{air} measured from approximately 2 \si{\meter} above ground level, yielding a "canopy layer" UHI (sUHI) - and T\textsubscript{surf} (sUHI) tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}. However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational aUHI studies suggests that that simplicity may not translate well to real world analysis. In a points based assessment of 190 aUHI studies published between 1950 and 2007, \citet{Stewart2011} found that nearly 50\% of studies are "scientifically indefensible". In his analysis, each study was given a "passing" or "failing" grade and ranked based on methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronisity. Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain. 89\% lacked meta-data altogether. These findings suggest a significant gap between conceptual simplicity and practical realities in UHI analysis.

@@

In contrast to urban T\textsubscript{air} and aUHI assessment, where measurement techniques have not undergone significant recent evolution and methodological problems arise in large part from conceptual shortcomings and simple carelessness, study of the urban effect on land T\textsubscript{surf} is relatively new and significantly more complex. As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases. In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid proliferation in study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of upwelling thermal infrared radiation (TIR). In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:

@@

Remote sensed study of the urban effect on land T\textsubscript{surf} has been critical in characterizing large-scale spatial and temporal patterns of sUHI \citep{Peng2012, Streutker2003, Imhoff2010}. However, the methodological and conceptual concerns raised in \citet{Stewart2011} in combination with the wide (and broadening) range of methodologies for remote sensed urban T\textsubscript{surf} retrieval should prompt critical assessment of the remote sensed sUHI literature. At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available. In addition, an increasingly dogmatic focus on improving spatial and radiometric resolutions has left other biases inherent in remote sensing of urban environments largely ignored in sUHI literature. This suggests that similar conclusions to those found of aUHI literature in \citet{Stewart2011} are possible (if not likely) for study of sUHI. Assessment of these biases is, therefore, paramount to ensure assessments of the urban effect on T\textsubscript{surf} are robust, methodologically valid, and comparable across techniques, space, and time.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,
	height=7in,
	keepaspectratio]{model}
	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width \textit{H}, height 2\textit{H}, and inter element spacing \textit{H}. Figure from \cite{Adderley2015}.}
	\label{model}
\end{figure}

>>>>

\chapter{Introduction}
\label{intro}

Urban development drastically changes the character of the surface. Replacement of natural terrain with roads, buildings, parks, and other urban features modifies the geometry of the surface as well as its thermal, radiative, moisture, and aerodynamic properties. These modifications, combined with anthropogenic emissions of heat, result in a distinct urban climate; one which is often warmer, a phenomenon termed the urban heat island effect (UHI). As both the Earth's population and the urbanized fraction of that population increase \citep{Nations2014}, cities grow and more people are exposed to urban-modified atmospheres. Thus, understanding how cities interact with the climate across spatiotemporal scales has important implications for human health, energy efficiency, and for informing future urban development. This, in part, has prompted significant expansion in study of the urban effect on climates, with particular focus on characterization of the urban effect on surface (T\textsubscript{surf}) and air (T\textsubscript{air}) temperatures.

@@

Study of the UHI with respect to T\textsubscript{surf} (sUHI) and T\textsubscript{air} (aUHI) - typically referring to T\textsubscript{air} measured from approximately 2 \si{\meter} above ground level, yielding a "canopy layer" UHI (clUHI) - tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}. However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational aUHI studies suggests that that simplicity may not translate well to real world analysis. In a points based assessment of 190 aUHI studies published between 1950 and 2007, \citet{Stewart2011} found that nearly 50\% of studies are "scientifically indefensible". In his analysis, each study was given a "passing" or "failing" grade and ranked based on methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronisity. Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain. 89\% lacked meta-data altogether. These findings suggest a significant gap between conceptual simplicity and practical realities in UHI analysis.

@@

In contrast to urban T\textsubscript{air} and aUHI assessment, where measurement techniques have not undergone significant recent evolution and methodological problems arise in large part from conceptual shortcomings and simple carelessness, study of the urban effect on land T\textsubscript{surf} is relatively new and significantly more complex. As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases. In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in remote sensing of thermal infrared radiation (TIR) upwelling from urban environments has lead to a rapid proliferation of study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of TIR. In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:

@@

Remote sensed study of the urban effect on land T\textsubscript{surf} has been critical in characterizing large-scale spatial and temporal patterns of sUHI \citep{Peng2012, Streutker2003, Imhoff2010}. However, the methodological and conceptual concerns raised in \citet{Stewart2011} in combination with the wide (and broadening) range of methodologies for remote sensed urban T\textsubscript{surf} retrieval should prompt critical assessment of the remote sensed sUHI literature. At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available and questions raised in critical assessments of remote sensed urban T\textsubscript{surf} in \cite{Roth1989} and \cite{Voogt2003} have gone largely unaddressed. An increasingly dogmatic focus on improving spatial and radiometric resolutions has left other biases inherent in remote sensing of urban environments largely ignored in sUHI literature. This suggests that similar conclusions to those found of aUHI literature in \citet{Stewart2011} are possible (if not likely) for study of sUHI. Assessment of these biases is, therefore, paramount to ensure assessments of the urban effect on T\textsubscript{surf} are robust, methodologically valid, and comparable across techniques, space, and time.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,
	height=7in,
	keepaspectratio]{model}
	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width \textit{H}, height 2\textit{H}, and inter element spacing \textit{H}. Figure from Adderley et al. (2015).}
	\label{model}
\end{figure}

### 2017-05-10 paper1.tex
Add files via upload

\begin{figure}[H]
	\includegraphics[width=\textwidth]{spectairtsurf}
	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \citep{Berk1987} for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
	\label{spectairtsurf}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=\textwidth]{dometrans}
	\caption{A comparison of dome transmittance for a Kipp \& Zonen silicone domed pyrgeometer (data supplied by Kipp \& Zonen \textit{pers. comm.}) and at-sensor radiance (T\textsubscript{surf} = 300 \si{K}, T\textsubscript{air} = 300 \si{K}, $z$ = 30 \si{m})  for "typical" and extended bandpasses.}
	\label{dometrans}
\end{figure}

@@

To save computational time and facilitate broader use of the method, a parameterization scheme was developed to simplify the correction process. In the parameterization, a suite of radiative transfer simulations is run prior to correction (as opposed to at each time step) to retrieve $\tau$ for a range of humidities and path lengths. T\textsubscript{surf} and T\textsubscript{air} are held constant in these simulations as they have a relatively small effect on $\tau$. Simulations for path lengths from 1 to 100 \si{\meter} and water vapor mass densities from 0.1 to 23.5 \si{\gram\per\meter\cubed} are included in Appendix \ref{appendixa}. Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged, surface-to-sensor path length calculated using the SUM \citep{Soux2004} and a simplified DBM. This path length represents the average distance from the sensor to the surface as "seen" by the sensor. For each time step, measured humidity is used to infer a hemispherical transmittance ($\overline{\tau_{\Phi}}$) for the approximated surface-sensor geometry. A parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) can then be calculated by using measured (in this case, profile averaged) T\textsubscript{air} to decompose measured L\textsubscript{up} into irradiance received by the sensor from the surface $L_{surf}^{at-sensor}$ and from the atmosphere $L_{atm}^{at-sensor}$ via,

>>>>

\begin{figure}[H]
	\includegraphics[width=\textwidth]{spectairtsurf}
	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 (Berk et al., 1987) for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
	\label{spectairtsurf}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=\textwidth]{dometrans}
	\caption{A comparison of dome transmittance for a Kipp \& Zonen silicone domed pyrgeometer (data supplied by Kipp \& Zonen, \textit{pers. comm.}) and at-sensor radiance (T\textsubscript{surf} = 300 \si{K}, T\textsubscript{air} = 300 \si{K}, $z$ = 30 \si{m})  for "typical" and extended bandpasses.}
	\label{dometrans}
\end{figure}

@@

To save computational time and facilitate broader use of the method, a parameterization scheme was developed to simplify the correction process. In the parameterization, a suite of radiative transfer simulations is run prior to correction (as opposed to at each time step) to retrieve $\tau$ for a range of humidities and path lengths. T\textsubscript{surf} and T\textsubscript{air} are held constant in these simulations as they have a relatively small effect on $\tau$. Simulations for path lengths from 1 to 48 \si{\meter} and water vapor mass densities from 0.1 to 23.5 \si{\gram\per\meter\cubed} are included in Appendix \ref{appendixa}. Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged, surface-to-sensor path length calculated using the SUM \citep{Soux2004} and a simplified DBM. This path length represents the average distance from the sensor to the surface as "seen" by the sensor. For each time step, measured humidity is used to infer a hemispherical transmittance ($\overline{\tau_{\Phi}}$) for the approximated surface-sensor geometry. A parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) can then be calculated by using measured (in this case, profile averaged) T\textsubscript{air} to decompose measured L\textsubscript{up} into irradiance received by the sensor from the surface $L_{surf}^{at-sensor}$ and from the atmosphere $L_{atm}^{at-sensor}$ via,
